Segregation and Diversity Through an Information Lens

### 2016-11-04 revision.tex
cleaning up

\title{An Information Lens on Demographic Diversity} % work on this
\author{\emph{Phil Chodrow}}
\date{\emph{\today}}

\usepackage{pc_writeup}
\usepackage{pc_math}
\usepackage[colorinlistoftodos]{todonotes}

@@

\section{Introduction} \label{sec:intro}
	\begin{itemize}
		\item While many measures are available for the measurement of spatial diversity, we are not aware of quantitative methods for \emph{identifying} spatial structure. 
		\item We endorse an information-theoretic approach to the study of diversity and segregation. In supporting this approach, we make two main contributions: 
		\begin{itemize}
			\item First, we argue that a simple suite of information measures -- entropy, mutual information, and local information -- intuitively measure global diversity, spatial unevenness, and spatial exposure. The mean local information is novel, and we develop some of its useful properties. This triplet of information measures readily lends itself to quantifying dimensions of segregation within and between cities. 
			\item Measures such as those we develop above can be viewed as measuring different kinds of sociospatial structure. To our knowledge, no such analysis methods have been proposed for \emph{identifying} that structure. Building on our information methods, we develop an algorithm for clustering urban areas into regions whose demographics are near constant in space. The result allows the analyst to both easily visualize spatial patterns of difference and conduct analysis using a natural partition of urban space. 
		\end{itemize}
	\end{itemize}
\section{Lit Review, Previous Work. Why our problem is new.} \label{sec:previous}
	Core points: 
	\begin{itemize}
		\item Previous work is missing unification
		\item Previous work is missing a practical component: how do you compute and learn about urban structure with it? 
	\end{itemize}

\section{Segregation and Information} \label{sec:information}
	Information measures quantify structure through an epistemic framework. The core insight is that systemic structure enables \emph{prediction}, while systemic variability can make such prediction difficult. Two systemic features are related when knowledge of one enables better prediction of the other.

	We can formalize the idea of prediction quality by considering the following ``guessing game.'' We discuss the weather, and I ask you to give me a \emph{predictive distribution} on the question of whether or not it will rain tomorrow; e.g. ``70\% yes, 30\% no.'' Call this predictive distribution $p$. Tomorrow evening I give you a reward $R(p,Z)$ based on your predictive distribution $p$ and the observed outcome $Z$. For our game to be a true ``guessing game,'' we might expect the reward function $R$ to possess the following intuitive properties: 
	\begin{enumerate}
		\item \textbf{Honesty:} The reward scheme $R$ encourages you to report your true beliefs about the rain.
		\item \textbf{Fairness:} Your reward $R(p,Z)$ you receive depends only on the observed outcome, not ``what could have happened.''
		\item \textbf{Smoothness:} Your reward increases smoothly as a function of the probability you assigned to the observed outcome. 
	\end{enumerate}
	It is possible to show (see e.g. \cite{Cover1991}) that these properties uniquely determine the logarithmic reward function $R(p,Z) = \log p(Z)$ up to an affine transformation. This logarithmic reward is the fundamental unit of information-theoretic analysis. It is customary to focus on the \emph{loss function} $L(p,Z) = - R(p,Z)$. Large loss corresponds to poor prediction. Systemic structure is present when knowledge of a system feature enables us to decrease the loss in such a guessing game.

	\begin{figure}
		\includegraphics[width=\textwidth]{figs/checkerboard.png}
		\caption{Information and the checkerboard}. \label{fig:checkerboard}
	\end{figure}

@@

		As an introductory example, consider the following guessing game: I pick a person at random from a city, and ask you to assign a predictive distribution on their race $Y$, where $Y$ takes possible values in the alphabet  $\mathcal{Y} = \{\text{Asian, Black, Hispanic, Other, White}\}$. There is a true distribution $p(Y)$ corresponding to the proportions of each race in the city in question. It is possible to show that your optimal guess is precisely this distribution $p$. The optimal expected loss is 
		\begin{equation}
			\E_Y[L(p,Y)] = -\E_Y[R(p,Y)] = -\E_Y[\log p(Y)] \triangleq H(Y), 
		\end{equation}
		also known as the (Shannon) \emph{entropy} of $p$.

		The Shannon entropy is a natural measure of the global diversity of a city. When $H(Y)$ is low, guessing a randomly-chosen individual's race is easy. An extreme example of this case is when $H(Y) = 0$; in this case, the true distribution $p(Y)$ is unity for one race and zero for the others; i.e. the city is monoracial. Figure \ref{fig:checkerboard}(a) gives an example of this case in a two-race city. Perfect prediction without loss is possible because this city lacks diversity. On the other hand, when $H(Y)$ is high, guessing a randomly-chosen individual's race is difficult. An extreme example of this case is when $H(Y)$ achieves its maximum value of $\log \abs{\mathcal{Y}}$, in which case each race is represented equally. The cities in Figure \ref{fig:checkerboard}(b)-(d) all have maximum demographic entropy. The underlying distribution gives you grounds on which to add probabilistic weight to one race over another; you may as well roll a fair (five-sided) die. Prediction is difficult because of the city's diversity.

@@

		The authors of \cite{Reardon2004} define spatial evenness as ``the extent to which groups are similarly distributed in residential space.'' \emph{Un}evenness, then, is the extent to which these groups are dissimilarly distributed in space. This physical description has an epistemic analogue: knowing where a resident lives conveys information about their race, in the sense of enabling better prediction. To quantify this intuition, consider two versions of the guessing game above. In the first version, I simply choose a random resident and ask you to guess as before. In the second, I choose a random resident but then tell you their residence. In the presence of unevenness, this will generally enable you to improve your predictions. Let $X$ be a random variable denoting the spatial residence. Then, your expected loss in the first game is still $H(Y)$, and your expected loss in the second is $H(Y|X) \triangleq -\E_{Y}[\log p(Y)|X]$. The \emph{mutual information} $I(X,Y)$ is the predictive improvement associated with my telling you the residence: 
		\begin{equation}
			 I(X,Y) \triangleq H(Y) - H(Y|X)\;.
		\end{equation}
		$H(Y)$ is the expected loss without any knowledge of $X$, while $H(Y|X)$ is the expected loss when you are provided the value of $X$ as well. Since knowing $X$ can only improve prediction, $I(X,Y) \geq 0$. Furthermore, it is possible to show that 
		\begin{equation}
			I(X,Y) = D[p(X,Y)\|p(X)p(Y)]\;,
		\end{equation}
		where 
		\begin{equation}
			D[q\|r] \triangleq \sum_a q(a) \log \frac{q(a)}{r(a)}
		\end{equation}
		is the Kullback-Leibler (KL) divergence of $r$ from $q$. Though not an axiomatic metric, the KL divergence may be viewed as a measure of distance between two distributions. The mutual information $I(X,Y)$ can therefore also be viewed as the distance of the true distribution $p(X,Y)$ from the hypothetical distribution $p(X)p(Y)$ in which the variables $X$ and $Y$ are independent. $I(X,Y)$ therefore measures the degree of dependence between these variables.

		The author of \cite{Roberto2015a} uses the mutual information $I(X,Y)$ -- which she refers to as the Divergence Index -- as a measure of (un)evenness. To see this within our epistemic framework, $I(X,Y) = 0$, $X$ and $Y$ are independent, indicating that knowing $X$ does not improve your ability to guess $Y$. City (b) in Figure \ref{fig:checkerboard} illustrates this situation; since ``everywhere is the same,'' location knowledge is of no predictive value. If we were playing in City (b), your performance in the second guessing game would then be equal to your performance in the first. Physically, this situation corresponds to perfect evenness: the distribution of races is $p(Y|X)$ is the same at every location $X$. In contrast, when $I(X,Y)$ achieves its maximum value of $H(Y)$, $X$ fully determines $Y$. This situation is illustrated by Figure \ref{fig:checkerboard}(c); if I tell you that a resident lives in the left- or right-hand side of the city, you can guess their race with perfect accuracy. This situation corresponds to one in which $H(Y|X) = 0$, i.e. every location is monoracial, reflecting maximal sociospatial unevenness.

@@

		The authors of \cite{Reardon2004} define \emph{spatial exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' We can construct an epistemic analogue of spatial exposure by considering the following pair of guessing games. In each, I pick a resident at random from a fixed ``local spatial environment'' that we both know. In the first game, I ask you to guess that resident's race directly. In the second, I tell you where \emph{within that environment} the resident lives. If the spatial environment in question is $R$, then your expected loss in the first game is $H(Y|X\in B)$ and in the second $H(Y|X, X\in B)$. The prediction improvement associated with knowing residence within this local environment is 
		\begin{equation}
			I(X,Y | X \in B) = H(Y|X\in B) - H(Y|X, X\in B)\;.
		\end{equation}
		Our core claim is that $I(X,Y|X \in B)$ -- the \emph{local (mutual) information} -- is a measure of spatial exposure \emph{in unevenly distributed populations}. This qualification is important, since in a perfectly even population such as Figure \ref{fig:checkerboard}(b),  $I(X,Y) = 0$ and $I(X,Y|X\in B) = 0$ as well. Consider, however, the two unevenly distributed populations in Figure \ref{fig:checkerboard} (c) and (d). In Figure (c), if we play our two guessing games at a locale entirely within the left- or righthand halfs, the population is locally homogeneous and you can make perfect predictions regardless of the specific residence. Only if we play the game on a locale straddling the middle boundary between the two groups does knowing a resident's specific residence help you to make predictions. If our locale $B$ perfectly straddles the boundary, so that the population proportions are $(\frac{1}{2},\frac{1}{2})$, then we have $H(Y|X\in B) = \log 2$ and $H(Y|X,X\in B) = 0$, so $I(X,Y|X \in B) = \log 2$.

		This example illustrates that the local information $I(X,Y|X\in B)$ is high when $B$ is an area of demographic transition. When two cities are similarly uneven (similar $I(X,Y)$), a city with more such transitional areas has more exposure between groups, since residents of one area are closer to demographically different residents of another area. To illustrate, the city of Figure \ref{fig:checkerboard}(d) has the same level of spatial unevenness as \ref{fig:checkerboard}(c), but greater spatial exposure.

		For practical data analysis, it is necessary to understand how the value of $I(X,Y|X \in B)$ depends on the choice of the locale $B$. In particular, we should expect that when using fine-grained data and fine-grained spatial locales, our estimates of $I(X,Y|X \in B)$ should converge, and not vary wildly as we increase the resolution of our analysis. It is possible to prove mathematically that this is the case. The full proof is given in the appendix, but the idea of the proof is illustrative of the nature of the local information. It begins by noting that the local information is zero on areas of constant demographic structure, but achieves high values in transitional areas. In this sense it behaves like a mathematical derivative. Making this precise requires that we postulate a metric space $M$ and a \emph{probability field} $p(Y|X)$, $X \in M$ with the following properties: 
		\begin{enumerate}
		 	\item $p(y|x) > 0$ for all $x \in M$
		 	\item $p(y|x)$ is differentiable as a function of $x$ for all $x \in M$. 
		\end{enumerate}
		The probability field $p(y|x)$ corresponds to an idealization of our data; it can be viewed as an interpolation or approximation to the true data using a smooth curve.

		Define $I_r(x_0)$ to be the local information in a locale of radius $r$ around the point $x_0$:
		\begin{equation}
			I_r(x_0) = I(X,Y| \norm{X - x_0}^2 \leq r^2). 
		\end{equation}
		We may view $r$ as the resolution of our analysis. Under the stated conditions, the following approximation relationship holds: 
		\begin{equation}
			\lim_{r \rightarrow 0} \frac{I_r(x_0)}{r^2} = \frac{1}{4} \text{tr } J_Y(x_0), \label{eq:fisher_approx}
		\end{equation}
		where $J_Y(x_0)$ is the \emph{Fisher information in $Y$ about $x_0$}. $J_Y(x_0)$ is a matrix defined by the formula
		\begin{equation}
			J_Y(x_0) \triangleq \E_Y\left[\nabla_x \log p(Y|x) \left(\nabla_x \log p(Y|x)\right)^T \right]\;. 
		\end{equation}
		The trace $\text{tr }J_Y(x_0)$ of $J_Y(x_0)$ is simply the sum of the diagonal entries of $J_Y(x_0)$.

		Equation \eqref{eq:fisher_approx} implies that, for small $r$, the local information (when normalized by $r^2$) is approximately a constant that is independent of $r$. Furthermore, that constant is related to the derivatives of the probability field $p(Y|X)$. We can make this relationship more explicit by noting that 
		\begin{equation}
			\text{tr }J_Y(x_0) = \E_Y\left[\norm{\nabla_x \log p(Y|x_0)}^2\right]\;, \label{eq:derivative}
		\end{equation}
		which together with \eqref{eq:fisher_approx} shows that $I_r(x_0)$ is proportional to the mean magnitude of the derivative of $p(Y|X)$ evaluated at $x = x_0$.

		In the idealized mathematical framework, we may define the \emph{mean local information} as the average value of the trace of the Fisher information: 
		\begin{equation}
			J(X,Y) \triangleq \E_X[\text{tr } J_Y(X)].
		\end{equation}
		Using the relationship \eqref{eq:derivative}, we can recognize $J(X,Y)$ as the total variation of $\log p(Y|X)$ with respect to the probability measure induced by $X$. Intuitively, this is the average ``curviness'' of the field $p(Y|X)$.

		It is simple to adopt this measure $J(X,Y)$ for practical data analysis. The procedure is as follows: 
		\begin{enumerate}
		 	\item Partition the city under analysis into spatial locales $\{B_i\}$.
		 	\item Compute the local information $I(X,Y|X \in B_i)$ for each $i$. 
		 	\item Average the results, weighted by the population of each $B_i$. 
		\end{enumerate} 
		The explicit formula is 
		\begin{equation}
			J(X,Y) = \sum_i p(X \in B_i) I(X,Y|X \in B_i). \label{eq:formula}
		\end{equation}

	\subsection{Illustrations}
		\subsubsection{Comparative Analysis: Detroit and Philadelphia}
			To illustrate these methods, we use them to compare the cities of Philadelphia and Detroit. Our data is blockgroup-level demographics from the American Community Survey \cite{CensusRace}. As we show, computing $H(Y)$, $I(X,Y)$, and $J(X,Y)$ together highlights similarities and differences between these cities. 
			\todo{We need to generate a nice summary plot here and then fill this section in.}
			\begin{figure}
				\caption{Spatial analysis of racial diversity in Detroit (left) and Philadelphia (right). Top: Percentage of black residents in Detroit (left) and Philadelphia (right). Middle: Hexgrid of spatial resolution 1km imposed over each city, shaded by the value of $I(X,Y | X \in B_i)$. Bottom: Estimated contours of constant $I(X,Y | X \in B_i)$; high values correspond to spaces of demographic transition. Very bottom: Numerical summary, including the three information measures.} \label{fig:detroit_philly}
			\end{figure}

		\subsubsection{Comparative Diversity Profiles}
			\todo{Is there a fun way to explore the density point? E.g. figure out people per ``neighborhood'' in some way? Think about what I would want the analysis here to be.}

			We can achieve a broader picture by comparing $I(X,Y)$ and $J(X,Y)$ across a range of American cities, as we do in Figure \ref{fig:mutual_fisher}. The two measures $I(X,Y)$ and $J(X,Y)$ provide an intuitive sorting of major cities. Cities with small numbers of large, monoracial regions such as Atlanta and Detroit are sorted toward the bottom right while cities with more complicated patchworks of demographic variation like Philadelphia and New York are sorted toward the top right.

>>>>

\title{Segregation and Diversity Through an Information Lens} % work on this
\author{\emph{Phil Chodrow}}
\date{\emph{\today}}

\usepackage{pc_writeup}
\usepackage{pc_math}
\usepackage{csvsimple}
\usepackage{subcaption}
\usepackage{csquotes}

\usepackage[colorinlistoftodos]{todonotes}

@@

\section{Introduction} \label{sec:intro}

	% Residential segregation is a persistent fact of modern American cities. Spatial location within cities has been proposed as a generative force of disparity in education and income, as well as mental and physical health \cite{Sharkey2014,DiezRoux2001,Dietz2002,Ludwig2013,Ludwig2012}. Residential segregation -- the differential distribution of different racial groups across urban space -- would then be itself a driver of further inequalities.

	The quantification of segregation has been a topic of lively conversation within sociology through the last seventy years. This conversation may be coarsely partitioned into two interlocking threads. The first thread concerns the \emph{conceptual dimensions} of segregation, with each dimension reflecting an aspect of the ways in which groups may be separated in space. Most modern authors prefer a subset of the dimensions proposed by Massey and Denton \cite{Massey1988}: evenness, exposure, concentration, centralization, and clustering. Recent writers \cite{Reardon2002,Reardon2004,Roberto2015a,Roberto2015} have tended to focus on the dimensions of evenness and exposure, with \cite{Reardon2004} arguing that the additional clustering and concentration may be considered special cases.

	The second thread concerns the appropriate methods with which to quantify these dimensions. Most fall into one of two broad classes of mathematical structure.\footnote{These categories are related to but not identical to the standard classification introduced by \cite{Coulter1989}.} \emph{Combinatorial} measures are the most ``physical:'' they measure the probability that two individuals selected under specifiable conditions will fall into two specified racial categories \cite{Goodman1954,James1986}. Combinatorial measures may therefore be intepreted as probabilities of encounters between members of racial groups. They may be differentiated according to the conditions of selection and the choice of categories. Contrasting to combinatorial measures are \emph{information} measures \cite{Theil1971,Theil1967,Roberto2015a}. Information measures are nonphysical in the sense that they do not correspond to probabilities of events. Rather, they reflect an epistemic viewpoint, measuring the ability of an observer to make guesses about individuals' identities when supplied with various kinds of information. Combinatorial and information measures do not exhaust all possibilities, however. For example, \cite{Wong1999} shows that an parametric approach based on the standard deviational ellipses of different groups may also be used to characterize city-scale segregation patterns, and \cite{Osullivan2007} expands this approach using nonparametric methods.

	Theorists have proposed multiple methods for each combination of dimension and mathematical framework, leading to a proliferation of segregation measures. While an abundance of options is in principle a boon to the analyst, relatively little guidance exists on which measures to use \emph{together}.\footnote{\cite{Massey1988} provides some guidance, but this is based on a statistical comparison of a variety measures on data, rather than from first principles.} Some authors use different mathematical paradigms for different dimensions; for example, \cite{Reardon2004} develops a combinatorial approach to evenness and an information theoretic approach to exposure. Considering the intrinsically multidimensional character of segregation itself, a mathematically coherent ensemble of measures is called for. This essay therefore advocates an approach to the study of residential segregation that is \emph{conceptually multidimensional}, \emph{computationally practical}, and \emph{mathematically unified}.

	\textbf{Conceptually multidimensional.} ``Segregation'' is not a single quantifiable number; it is an intrinsically multidimensional phenomenon that therefore requires a coordinated ensemble of distinct measures. As far as possible, each measure should correspond to exactly one dimension of segregation. Furthermore, each measure should be interpreted in the context of each of the others. In this essay, we adopt the categories of evenness and exposure as emphasized by \cite{Reardon2004}, to which we add the dimension of global diversity of the city as a whole.

	\textbf{Computationally practical.} Writing in 2004, the authors of \cite{Reardon2004} observe that 
	\begin{displayquote}

		[a]t present, few of the proposed spatial segregation measures have been used in published empirical segregation research. These measures have been ignored in part because they typically are more difficult to compute than the aspatial measures.

	\end{displayquote}
	Over a decade later, computing resources have expanded dramatically in capability and accessibility. While our focus in this essay is theoretical, our ultimate goal is to place effective tools in the hands of practitioners. To this end, we emphasize interpretability, performance, and ease-of-use in the methods we develop. All are freely available as a package for the \texttt{R} programming language, as described in the Appendix.

	\textbf{Mathematically unified.} In the measurement of segregation, we typically treat a city as a collection of different types of individuals, distributed differentially in space. The distinct dimensions of segregation correspond to the distinct questions we can ask about these distributions. Information theory is the mathematical framework most suited for the comparison of distributions. Rather than developing  measures for each dimension separately, we show how that an information-theoretic framework leads naturally to an ensemble of conceptually coherent, interpretable metrics resting on a shared mathematical foundation.

	Our original contribution to this program has three components. The first is a novel measure of spatial exposure -- the local information -- and prove some of its properties. As we show, the local information distinguishes cities with starkly separated monoracial regions from cities with more complex patchwork patterns of racial difference. We argue that the local information, together with global entropy and mutual information, constitute a coherent ensemble of measures for the multidimensional quantification of residential segregation. The second is a novel, information-based algorithm for identifying the spatial structure of residential segregation. The algorithm clusters urban areas into regions whose demographics are relatively constant in space, allowing the analyst to study organic, demographically coherent regions rather than the partially-arbitrary tract or blockgroup demarcations provided by the U.S. Census. The third contribution is to share the tools we develop freely with practitioners: our suite of information measures and our clustering tools are available as open-source software for the \texttt{R} statistical programming language.

\section{Information and Segregation} \label{sec:information}
	Information theory has long been a mainstay of engineering, computer science, and the natural sciences. Somewhat more recently, information theory has found application in the study of urban phenomena. Theorists have used information measures to have been used to estimate the distribution of populations and resources in space \cite{Webber1979}, guide urban zoning decisions \cite{Batty1972}, identify spatial hierarchies in population distributions \cite{Batty1974,Batty1976}, and measure the complexity of urban phenomena \cite{Bettencourt2015,Batty2014a}. In this section, we develop some key information theoretic concepts and relate them to the dimensions of residential segregation.

	Information theory quantifies structure through an epistemic framework. The core insight is that systemic structure enables \emph{prediction}, while systemic variability can make such prediction difficult. Two systemic features are related when knowledge of one enables better prediction of the other. We can formalize the idea of prediction by considering the following ``guessing game.'' We discuss the weather, and I ask you to give me a \emph{predictive distribution} on the question of whether or not it will rain tomorrow. For example, you might tell me your predictive distribution $p$, where $p(\text{Rain}) = 70\%$ $p(\text{No Rain}) = 30\%$. Tomorrow evening I give you a reward $R(p,W)$ based on your predictive distribution $p$ and the observed weather $W$. For this game to be a true ``guessing game,'' we would expect the function $R$ by which I reward your guess to possess the following intuitive properties: 
	\begin{enumerate}
		\item \textbf{Honesty:} The reward function $R$ encourages you to report your true beliefs about the rain.
		\item \textbf{Fairness:} The reward $R(p,W)$ you receive depends only on your prediction $p$ and the observed outcome, not ``what could have happened.''\footnote{An ``honest'' and ``fair'' reward function is more standardly called ``proper'' and ``local'' among information theorists. 
		}
		\item \textbf{Smoothness:} Your reward increases smoothly as a function of the probability you assigned to the observed outcome. 
	\end{enumerate}
	It is possible to show that these properties uniquely determine the logarithmic reward function $R(p,W) = \log p(W)$ up to an affine transformation.\footnote{\cite{Cover1991,Csiszzr2004}} This logarithmic reward is therefore a fundamental unit of information-theoretic analysis. It is traditional to focus on the negative of the reward function -- the \emph{loss function} $L(p,W) = - R(p,W)$. Large loss reflects to poor prediction. Systemic structure is present when knowledge of a system feature enables us to decrease the loss in such a guessing game.

@@

		Consider the following predictive guessing game. You and I are jointly studying residential segregation in Atlanta. I pick a person at random from the city, and ask you to give predictive distribution on their race $Y$, where $Y$ takes possible values in the alphabet  $\mathcal{Y} = \{\text{Asian, Black, Hispanic, Other, White}\}$. There is a true distribution $p(Y)$ corresponding to the proportions of each race in the city in question; in Atlanta in 2014 this true distribution is 
		\begin{align*}
			p(\text{Asian})    &= 6\% \quad
			p(\text{Black})    = 47\% \quad
			p(\text{Hispanic}) = 9\% \\
			p(\text{Other})    &= 4\% \quad
			p(\text{White})    = 35\%\;. 
		\end{align*}
		It is possible to show that your optimal guess in our game is precisely this distribution $p$. Your optimal expected loss is therefore
		\begin{equation*}
			\E_Y[L(p,Y)] = -\E_Y[R(p,Y)] = -\E_Y[\log p(Y)] \triangleq H(Y), 
		\end{equation*}
		where $\E_Y$ is the expectation or averaging operator with respect to $Y$. $H(Y)$ is the Shannon \emph{entropy} of $p$ \cite{Shannon1948}. Throughout this paper logarithms will use the natural base $e = 2.718\ldots$, in which case all information measures have dimensionless units of nats.

		\begin{figure}
			\includegraphics[width=\textwidth]{figs/checkerboard.pdf}
			\caption{Schematic cities illustrating three dimensions of segregation. City (a) lacks global diversity, while Cities (b)-(d) are maximally globally diverse. City (b) is perfectly spatially even, whereas Cities (c) and (d) are maximally uneven. City (d) has higher spatial exposure than City (c) due to more contact points between racially distinct neighborhoods.}. \label{fig:checkerboard}
		\end{figure}

		The Shannon entropy is a natural measure of the global diversity of a city. When $H(Y)$ is low, guessing a randomly-chosen individual's race is easy. An extreme example of this case is when $H(Y) = 0$; in this case, the true distribution $p(Y)$ is unity for one race and zero for the others; i.e. the city is monoracial. Figure \ref{fig:checkerboard}(a) illustrates this case in a schematic two-race city. Perfect prediction without loss is possible because this city lacks diversity. On the other hand, when $H(Y)$ is high, guessing a randomly-chosen individual's race is difficult. An extreme example of this case is when $H(Y)$ achieves its maximum value of $\log \abs{\mathcal{Y}}$, in which case each race is represented equally. The cities in Figure \ref{fig:checkerboard}(b)-(d) all have maximum demographic entropy. The underlying distribution gives you no grounds on which to add probabilistic weight to one race over another; flip a fair coin. Prediction is difficult because of the city's diversity.

@@

		The authors of \cite{Reardon2004} define spatial evenness as ``the extent to which groups are similarly distributed in residential space.'' \emph{Un}evenness, then, is the extent to which these groups are dissimilarly distributed in space. This physical description has an epistemic analogue. In an uneven city, knowing where a resident lives conveys information about their race, in the sense of enabling better prediction. To quantify this intuition, consider two versions of the guessing game in Atlanta. In the first version, I simply choose a random resident and ask you to guess as before. In the second, I choose a random resident but then tell you their location of residence. In the presence of unevenness, this will generally enable you to improve your predictions. For example, in Atlanta knowing that the resident lives in the predominantly white northern part of the city may substantially change your predictive distribution. We can quantify the value of this information as follows. Let $X$ be a random variable denoting the spatial residence. Then, your expected loss in the first game is still $H(Y)$, and your expected loss in the second is $H(Y|X) \triangleq -\E_{Y}[\log p(Y)|X]$. The \emph{mutual information} $I(X,Y)$ is the predictive improvement associated with my telling you the residence: 
		\begin{equation*}
			 I(X,Y) \triangleq H(Y) - H(Y|X)\;.
		\end{equation*}
		Since knowing $X$ can only improve prediction, $I(X,Y) \geq 0$. Furthermore, it is possible to show that 
		\begin{equation}
			I(X,Y) = D[p(X,Y)\|p(X)p(Y)]\;, \label{eq:info_and_divergence}
		\end{equation}
		where 
		\begin{equation*}
			D[q\|r] \triangleq \sum_a q(a) \log \frac{q(a)}{r(a)}
		\end{equation*}
		is the Kullback-Leibler (KL) divergence of $r$ from $q$. Though not an axiomatic metric, the KL divergence may be viewed as a measure of distance between two distributions. Equation \eqref{eq:info_and_divergence} thus shows that the mutual information $I(X,Y)$ can be viewed as the distance of the true joint distribution $p(X,Y)$ of location and race from the hypothetical distribution $p(X)p(Y)$ in which location and race are independent. $I(X,Y)$ therefore measures the degree of dependence between these variables. In this sense, it resembles a correlation coefficient, but has the stronger property that $I(X,Y) = 0$ implies true statistical independence.

		The mutual information can be viewed as a measure of spatial unevenness (\cite{Roberto2015a}). To see why the mutual information should measure unevenness, consider first the case in which $I(X,Y) = 0$. Since $X$ and $Y$ are independent, knowledge of $X$ does not improve your ability to guess $Y$. City (b) in Figure \ref{fig:checkerboard} illustrates this situation; since ``everywhere is the same,'' location knowledge is of no predictive value. If we were playing in City (b), your performance in the second guessing game would then be equal to your performance in the first. Physically, this situation corresponds to perfect evenness: the distribution of races is $p(Y|X)$ is the same at every location $X$. In contrast, when $I(X,Y)$ achieves its maximum value of $H(Y)$, $X$ fully determines $Y$. This case is illustrated by Figure \ref{fig:checkerboard}(c); if I tell you that a resident lives in the left- or right-hand side of the city, you can guess their race with perfect accuracy. Every location is monoracial, reflecting maximal sociospatial unevenness.

@@

		The authors of \cite{Reardon2004} define \emph{spatial exposure} as ``the extent that members of one group encounter members of another group...in their local spatial environments.'' We can construct an epistemic analogue of spatial exposure by considering the following pair of guessing games in Atlanta. In each, I pick a resident at random from a small neighborhood that we both know in advance. In the first game, I then ask you to guess that resident's race immediately. In the second, I tell you where \emph{within that neighborhood} the resident lives. If the neighborhood in question is $B$, then your expected loss in the first game is $H(Y|X\in B)$ and in the second $H(Y|X, X\in B)$. The predictive improvement associated with knowing residence within this neighborhood is therefore
		\begin{equation*}
			I(X,Y | X \in B) = H(Y|X\in B) - H(Y|X, X\in B)\;.
		\end{equation*}
		The measure $I(X,Y|X \in B)$ -- the \emph{local (mutual) information} -- is a measure of spatial exposure \emph{in unevenly distributed populations}. This qualification is important, since in a perfectly even population such as Figure \ref{fig:checkerboard}(b),  $I(X,Y) = 0$ and $I(X,Y|X\in B) = 0$ as well. Consider, however, the two unevenly distributed populations in Figure \ref{fig:checkerboard} (c) and (d). In \ref{fig:checkerboard}(c), if we play our two guessing games at a locale entirely within the left- or right-hand halves, the population is locally homogeneous and you can make perfect predictions regardless of the specific residence. Only if we play the game on a locale straddling the middle boundary between the two groups does knowing a resident's specific residence help you to make predictions. If $B$ perfectly straddles the boundary, so that the population proportions are $(\frac{1}{2},\frac{1}{2})$, then we have $H(Y|X\in B) = \log 2$ and $H(Y|X,X\in B) = 0$, so $I(X,Y|X \in B) = \log 2$.

		This example illustrates that the local information $I(X,Y|X\in B)$ is high when $B$ is an area of demographic transition. When two cities are similarly uneven (have similar $I(X,Y)$), a city with more such transitional areas has more exposure between groups, since residents of one area are closer to demographically different residents of another area. To illustrate, the city of Figure \ref{fig:checkerboard}(d) has the same level of spatial unevenness as \ref{fig:checkerboard}(c), but greater spatial exposure.

		For practical data analysis, it is necessary to understand how the value of $I(X,Y|X \in B)$ depends on the choice of the locale $B$. In particular, we should expect that when using fine-grained data and fine-grained spatial locales, our estimates of $I(X,Y|X \in B)$ should converge, and not vary wildly as we increase the resolution of our analysis. It is possible to prove mathematically that this is indeed the case. The full proof is given in the appendix, but the idea of the proof is illustrative of the nature of the local information. It begins by noting that the local information is zero on areas of constant demographic structure, but achieves high values in transitional areas. In this sense it behaves like a mathematical derivative. Making this precise requires that we postulate a metric space $M$ and a \emph{probability field} $p(Y|X)$, $X \in M$ with the following properties: 
		\begin{enumerate}
		 	\item $p(y|x) > 0$ for all $x \in M$
		 	\item $p(y|x)$ is differentiable as a function of $x$ for all $x \in M$ and $y \in \mathcal{Y}$. 
		\end{enumerate}
		The probability field $p(y|x)$ corresponds to an idealization of our data; it can be viewed as an interpolation or approximation to the true data using a smooth function.

		Define $I_r(x)$ to be the local information in a locale of radius $r$ around the point $x$:
		\begin{equation*}
			I_r(x) = I(X,Y| \norm{X - x}^2 \leq r^2). 
		\end{equation*}
		We may view $r$ as the resolution of our analysis. Under the stated conditions, the following approximation relationship holds: 
		\begin{equation}
			\lim_{r \rightarrow 0} \frac{I_r(x)}{r^2} = \frac{1}{4} \text{tr } J_Y(x), \label{eq:fisher_approx}
		\end{equation}
		where $J_Y(x)$ is the \emph{Fisher information in $Y$ about $x$}. It is a matrix defined by the formula
		\begin{equation*}
			J_Y(x) \triangleq \E_Y\left[\nabla_x \log p(Y|x) \left(\nabla_x \log p(Y|x)\right)^T \right]\;. 
		\end{equation*}
		The trace $\text{tr }J_Y(x)$ is the sum of the diagonal entries of $J_Y(x)$.

		Equation \eqref{eq:fisher_approx} implies that, for small $r$, the local information in a neighborhood of radius $r$, when normalized by $r^2$, is approximately a constant that is independent of $r$. Furthermore, that constant is related to the derivatives of the probability field $p(Y|X)$. We can make this relationship more explicit by noting that 
		\begin{equation}
			\text{tr }J_Y(x) = \E_Y\left[\norm{\nabla_x \log p(Y|x)}^2\right]\;, \label{eq:derivative}
		\end{equation}
		which together with \eqref{eq:fisher_approx} shows that $I_r(x)$ is proportional to the mean magnitude of the derivative of $p(Y|X)$ evaluated at $X = x$.

		In the idealized mathematical framework, we may define the \emph{mean local information} as the average value of the trace of the Fisher information: 
		\begin{equation*}
			J(X,Y) \triangleq \E_X[\text{tr } J_Y(X)].
		\end{equation*}
		Using the relationship \eqref{eq:derivative}, we can recognize $J(X,Y)$ as the total variation of $\log p(Y|X)$ with respect to the probability measure induced by $X$. Intuitively, this is the average ``curviness'' of the field $p(Y|X)$.

		The measure $J(X,Y)$ may be easily adapted for practical computation. The procedure is as follows:
		\begin{enumerate}
		 	\item Partition the city under analysis into spatial locales $\{B_i\}$.
		 	\item Compute the local information $I(X,Y|X \in B_i)$ for each $i$. 
		 	\item Average the results, weighted by the population of each $B_i$. 
		\end{enumerate} 
		The explicit formula is 
		\begin{align}
			J(X,Y) &= \E_B[I(X,Y|X \in B)] \\ 
			       &= \sum_i p(X \in B_i) I(X,Y|X \in B_i). \label{eq:formula}
		\end{align}
		While the result \eqref{eq:fisher_approx} guarantees limiting resolution-independence in a mathematically idealized setting, in practice it is advisable to use locales $B_i$ of constant size and shape for all regions of analysis.
